Determining the alternatives for scalar implicature

### 2016-02-02 Paper/Paper.Rmd
incorporated some of Mika's comments

abstract: 
    Successful communication regularly requires listeners to make pragmatic inferences - enrichments beyond the literal meaning of a speaker's utterance. For example, when interpreting a sentence such as "Alice ate some of the cookies," listeners routinely infer that Alice did not eat all of them. A Gricean account of this phenomena assumes the presence of alternatives (like "all of the cookies") with varying degrees of informativity, but it remains an open question precisely what these alternatives are. To address this question, we collect empirical measurements of speaker and listener judgments about varying sets of alternatives across a range of scales and use these as inputs to a computational model of pragmatic inference. This approach allows us to test hypotheses about how well different sets of alternatives predict pragmatic judgments by people. Our findings suggest that comprehenders likely consider a broader set of alternatives beyond those logically entailed by the initial message.

@@

In our current work, we pick up where these previous studies left off,  considering the set of alternatives for implicature using the RSA model. To gain empirical traction on this issue, however, we broaden the set of scales we consider. Our inspiration for this move comes from work by @vantiel2014b, who examined a phenomenon that they dubbed "scalar diversity," namely the substantial difference in the strength of scalar implicature across a variety of scalar pairs (e.g., *liked* / *loved,* *palatable* / *delicious*). Making use of this diversity allows us to investigate the ways that different alternative sets give rise to implicatures of different strengths across scales.

@@

We recruited 30 workers on AMT. All participants were native English speakers and naive to the purpose of the experiment.

@@

Participants were presented with a target scalar item from five scales, selected from @vantiel2014b (see Figure \ref{fig:allScalesTable}). These were embedded in a sentence such as, "In a recent restaurant review someone said they thought the the food was \____," with a target scalar presented in the blank. Participants were then asked to generate plausible alternatives by responding to the question, "If they'd felt differently about the food, what other words could they have used instead of \____?" They were prompted to generate three unique alternatives.

@@

Conditions were run sequentially. In each condition we recruited 30 participants from Amazon Mechanical Turk (AMT). In the two alternatives condition, `r 30 - length(unique(exp1a$workerid))` participants were excluded for either failing to pass two training trials or because they were not native English speakers, leaving a total sample of `r length(unique(exp1a$workerid))` participants.\footnote{The majority of respondent data excluded from the two alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in later conditions.} In the four alternative condition,  `r 30 - length(unique(exp1b$workerid))` participants were excluded for either failing to pass two training trials or not being native English speakers, leaving a total sample of `r length(unique(exp1b$workerid))` participants. In the five alternative condition, `r 30 - length(unique(exp1c$workerid))` participants were excluded for either failing to pass two training trials or not being native English speakers, leaving a total sample of `r length(unique(exp1c$workerid))` participants.

@@

Figure \ref{fig:exp1Plots} plots estimated literal listener semantics for the three conditions. Each row shows a unique scalar family with items ordered horizontally by valence. Several trends are visible. First, in each scale, the alternatives spanned the star scale - there were scalar items that were highly compatible with both the lowest and highest numbers of stars. Second, participant compatibility judgments match intuitive literal semantics. That is, both weaker (e.g. *some* or *good*) and stronger (e.g. *all* or *excellent*) entailment items were seen as compatible with five-star ratings. This means participants' compatibility judgements reflect literal semantic intuitions, not enriched pragmatic judgments. We see clear variability between scalar families (literal semantic distributions for *memorable* are considerably different from *palatable*), but also substantial consistency across conditions.

@@

Model fit with human judgments was significantly improved by the inclusion of alternatives beyond the entailment items (Table 1). The two alternatives model contained only entailment items, which, under classic accounts, should be sufficient to generate implicature, but fit to data was quite poor with these items. The addition of a generic negative element produced some gains in performance, but much higher performance was found when we included four and five alternatives, with the alternatives derived empirically for the specific scale we used. An example fit for the five alternatives model is shown in Figure \ref{fig:fiveAltsScatter}.

>>>>

abstract: 
    Successful communication regularly requires listeners to make pragmatic inferences --- enrichments beyond the literal meaning of a speaker's utterance. For example, when interpreting a sentence such as "Alice ate some of the cookies," listeners routinely infer that Alice did not eat all of them. A Gricean account of this phenomena assumes the presence of alternatives (like "all of the cookies") with varying degrees of informativity, but it remains an open question precisely what these alternatives are. To address this question, we collect empirical measurements of speaker and listener judgments about varying sets of alternatives across a range of scales and use these as inputs to a computational model of pragmatic inference. This approach allows us to test hypotheses about how well different sets of alternatives predict pragmatic judgments by people. Our findings suggest that comprehenders likely consider a broader set of alternatives beyond those logically entailed by the initial message.

@@

In our current work, we pick up where these previous studies left off,  considering the set of alternatives for implicature using the RSA model. To gain empirical traction on this issue, we broaden the set of scales we consider. Our inspiration for this move comes from work by @vantiel2014b, who examined a phenomenon that they dubbed "scalar diversity," namely the substantial difference in the strength of scalar implicature across a variety of scalar pairs (e.g., *liked* / *loved,* *palatable* / *delicious*). Making use of this diversity allows us to investigate the ways that different alternative sets give rise to implicatures of different strengths across scales.

@@

We recruited 30 workers on Amazon Mechanical Turk (AMT). All participants were native English speakers and naive to the purpose of the experiment.

@@

Participants were presented with a target scalar item from five scales, selected from @vantiel2014b. These were embedded in a sentence such as, "In a recent restaurant review someone said they thought the the food was \____," with a target scalar presented in the blank. Participants were then asked to generate plausible alternatives by responding to the question, "If they'd felt differently about the food, what other words could they have used instead of \____?" They were prompted to generate three unique alternatives.

@@

Conditions were run sequentially. In each condition we recruited 30 participants from AMT. In the two alternatives condition, `r 30 - length(unique(exp1a$workerid))` participants were excluded for either failing to pass two training trials or because they were not native English speakers, leaving a total sample of `r length(unique(exp1a$workerid))` participants.\footnote{The majority of respondent data excluded from the two alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in later conditions.} In the four alternative condition,  `r 30 - length(unique(exp1b$workerid))` participants were excluded for either failing to pass two training trials or not being native English speakers, leaving a total sample of `r length(unique(exp1b$workerid))` participants. In the five alternative condition, `r 30 - length(unique(exp1c$workerid))` participants were excluded for either failing to pass two training trials or not being native English speakers, leaving a total sample of `r length(unique(exp1c$workerid))` participants.

@@

Figure \ref{fig:exp1Plots} plots estimated literal listener semantics for the three conditions. Each row shows a unique scalar family with items ordered horizontally by valence. Several trends are visible. First, in each scale, the alternatives spanned the star scale --- there were scalar items that were highly compatible with both the lowest and highest numbers of stars. Second, participant compatibility judgments match intuitive literal semantics. That is, both weaker (e.g. *some* or *good*) and stronger (e.g. *all* or *excellent*) entailment items were seen as compatible with five-star ratings. This means participants' compatibility judgements reflect literal semantic intuitions, not enriched pragmatic judgments. We see clear variability between scalar families (literal semantic judgments for *memorable* are considerably different from *palatable*), but also substantial consistency for individual scalar items across conditions (*good* in the two alternative condition is similar to *good* in the four and five alternative condition).

@@

Model fit with human judgments was significantly improved by the inclusion of alternatives beyond the entailment items (Table 1). The two alternatives model contained only entailment items, which, under classic accounts, should be sufficient to generate implicature, but fit to data was quite poor with these items. The addition of a generic negative element produced some gains, but much higher performance was found when we included four and five alternatives, with the alternatives derived empirically for the specific scale we used. An example fit for the five alternatives model is shown in Figure \ref{fig:fiveAltsScatter}.

### 2016-02-02 Paper/Paper.tex
incorporated some of Mika's comments

\begin{abstract}
Successful communication regularly requires listeners to make pragmatic
inferences - enrichments beyond the literal meaning of a speaker's
utterance. For example, when interpreting a sentence such as ``Alice ate
some of the cookies,'' listeners routinely infer that Alice did not eat
all of them. A Gricean account of this phenomena assumes the presence of
alternatives (like ``all of the cookies'') with varying degrees of
informativity, but it remains an open question precisely what these
alternatives are. To address this question, we collect empirical
measurements of speaker and listener judgments about varying sets of
alternatives across a range of scales and use these as inputs to a
computational model of pragmatic inference. This approach allows us to
test hypotheses about how well different sets of alternatives predict
pragmatic judgments by people. Our findings suggest that comprehenders
likely consider a broader set of alternatives beyond those logically
entailed by the initial message.

@@

In our current work, we pick up where these previous studies left off,
considering the set of alternatives for implicature using the RSA model.
To gain empirical traction on this issue, however, we broaden the set of
scales we consider. Our inspiration for this move comes from work by
{van Tiel}, Van Miltenburg, Zevakhina, \& Geurts (2014), who examined a
phenomenon that they dubbed ``scalar diversity,'' namely the substantial
difference in the strength of scalar implicature across a variety of
scalar pairs (e.g., \emph{liked} / \emph{loved,} \emph{palatable} /
\emph{delicious}). Making use of this diversity allows us to investigate
the ways that different alternative sets give rise to implicatures of
different strengths across scales.

@@

Participants were presented with a target scalar item from five scales,
selected from {van Tiel} et al. (2014) (see Figure
\ref{fig:allScalesTable}). These were embedded in a sentence such as,
``In a recent restaurant review someone said they thought the the food
was \_\_\_\_," with a target scalar presented in the blank. Participants
were then asked to generate plausible alternatives by responding to the
question, ``If they'd felt differently about the food, what other words
could they have used instead of \_\_\_\_?" They were prompted to
generate three unique alternatives.

@@

Conditions were run sequentially. In each condition we recruited 30
participants from Amazon Mechanical Turk (AMT). In the two alternatives
condition, 16 participants were excluded for either failing to pass two
training trials or because they were not native English speakers,
leaving a total sample of 14
participants.\footnote{The majority of respondent data excluded from the two alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in later conditions.}
In the four alternative condition, 7 participants were excluded for
either failing to pass two training trials or not being native English
speakers, leaving a total sample of 23 participants. In the five
alternative condition, 3 participants were excluded for either failing
to pass two training trials or not being native English speakers,
leaving a total sample of 27 participants.

@@

Figure \ref{fig:exp1Plots} plots estimated literal listener semantics
for the three conditions. Each row shows a unique scalar family with
items ordered horizontally by valence. Several trends are visible.
First, in each scale, the alternatives spanned the star scale - there
were scalar items that were highly compatible with both the lowest and
highest numbers of stars. Second, participant compatibility judgments
match intuitive literal semantics. That is, both weaker (e.g.
\emph{some} or \emph{good}) and stronger (e.g. \emph{all} or
\emph{excellent}) entailment items were seen as compatible with
five-star ratings. This means participants' compatibility judgements
reflect literal semantic intuitions, not enriched pragmatic judgments.
We see clear variability between scalar families (literal semantic
distributions for \emph{memorable} are considerably different from
\emph{palatable}), but also substantial consistency across conditions.

>>>>

\begin{abstract}
Successful communication regularly requires listeners to make pragmatic
inferences --- enrichments beyond the literal meaning of a speaker's
utterance. For example, when interpreting a sentence such as ``Alice ate
some of the cookies,'' listeners routinely infer that Alice did not eat
all of them. A Gricean account of this phenomena assumes the presence of
alternatives (like ``all of the cookies'') with varying degrees of
informativity, but it remains an open question precisely what these
alternatives are. To address this question, we collect empirical
measurements of speaker and listener judgments about varying sets of
alternatives across a range of scales and use these as inputs to a
computational model of pragmatic inference. This approach allows us to
test hypotheses about how well different sets of alternatives predict
pragmatic judgments by people. Our findings suggest that comprehenders
likely consider a broader set of alternatives beyond those logically
entailed by the initial message.

@@

In our current work, we pick up where these previous studies left off,
considering the set of alternatives for implicature using the RSA model.
To gain empirical traction on this issue, we broaden the set of scales
we consider. Our inspiration for this move comes from work by {van
Tiel}, Van Miltenburg, Zevakhina, \& Geurts (2014), who examined a
phenomenon that they dubbed ``scalar diversity,'' namely the substantial
difference in the strength of scalar implicature across a variety of
scalar pairs (e.g., \emph{liked} / \emph{loved,} \emph{palatable} /
\emph{delicious}). Making use of this diversity allows us to investigate
the ways that different alternative sets give rise to implicatures of
different strengths across scales.

@@

Participants were presented with a target scalar item from five scales,
selected from {van Tiel} et al. (2014). These were embedded in a
sentence such as, ``In a recent restaurant review someone said they
thought the the food was \_\_\_\_," with a target scalar presented in
the blank. Participants were then asked to generate plausible
alternatives by responding to the question, ``If they'd felt differently
about the food, what other words could they have used instead of
\_\_\_\_?" They were prompted to generate three unique alternatives.

@@

Conditions were run sequentially. In each condition we recruited 30
participants from AMT. In the two alternatives condition, 16
participants were excluded for either failing to pass two training
trials or because they were not native English speakers, leaving a total
sample of 14
participants.\footnote{The majority of respondent data excluded from the two alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in later conditions.}
In the four alternative condition, 7 participants were excluded for
either failing to pass two training trials or not being native English
speakers, leaving a total sample of 23 participants. In the five
alternative condition, 3 participants were excluded for either failing
to pass two training trials or not being native English speakers,
leaving a total sample of 27 participants.

@@

Figure \ref{fig:exp1Plots} plots estimated literal listener semantics
for the three conditions. Each row shows a unique scalar family with
items ordered horizontally by valence. Several trends are visible.
First, in each scale, the alternatives spanned the star scale --- there
were scalar items that were highly compatible with both the lowest and
highest numbers of stars. Second, participant compatibility judgments
match intuitive literal semantics. That is, both weaker (e.g.
\emph{some} or \emph{good}) and stronger (e.g. \emph{all} or
\emph{excellent}) entailment items were seen as compatible with
five-star ratings. This means participants' compatibility judgements
reflect literal semantic intuitions, not enriched pragmatic judgments.
We see clear variability between scalar families (literal semantic
judgments for \emph{memorable} are considerably different from
\emph{palatable}), but also substantial consistency for individual
scalar items across conditions (\emph{good} in the two alternative
condition is similar to \emph{good} in the four and five alternative
condition).
